A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-05-20 paper/paper.tex
fixed table
adding pareto plots
adding DOE in acknowledgments

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{url}
\usepackage{footmisc}
\usepackage{multicol}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,snakes,backgrounds}
\usetikzlibrary{mindmap,trees}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{plotmarks}

@@

\begin{document}

@@

\author{N. Neveu\footnotemark[7]\ \footnotemark[8]
  \and Y. Ineichen\footnotemark[2]\ \footnotemark[3]\ \footnotemark[4]
  \and A. Adelmann\footnotemark[2]\ \footnotemark[6]
  \and P. Arbenz\footnotemark[4]
  \and C. Bekas\footnotemark[3]
  \and A. Curioni\footnotemark[3]
  \and A. Kolano\footnotemark[2]\ \footnotemark[5]
  \and C. Kraus
}

@@

The authors thank the \textsc{SwissFEL} and AWA teams for contributing to the
  formulation of optimization problems. We gratefully acknowledge the computing resources provided on Bebop, a HPC cluster operated by the LCRC at ANL.
\appendix

>>>>

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{url}
\usepackage{footmisc}
\usepackage{multicol}
\usepackage{color}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,snakes,backgrounds}
\usetikzlibrary{mindmap,trees}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{plotmarks}
\usepackage{subcaption}

@@

\usepackage{array}
\newcolumntype{C}{ >{\centering\arraybackslash} m{2.5cm} }
\newcolumntype{D}{ >{\centering\arraybackslash} m{3cm} }

\newcommand{\nrnote}[1]{\textsf{{\color{violet}{ NN note:}   #1 }}}
\begin{document}

@@

\author{N. Neveu\footnotemark[7]\ \footnotemark[8]
  \and Y. Ineichen\footnotemark[2]\ \footnotemark[3]\ \footnotemark[4]
  \and A. Adelmann\footnotemark[2]\ \footnotemark[6]
  \and P. Arbenz\footnotemark[4]
  \and C. Bekas\footnotemark[3]
  \and A. Curioni\footnotemark[3]
  \and A. Kolano\footnotemark[2]\ \footnotemark[5]
  \and C. Metzger-Kraus 
}

@@

The authors thank the \textsc{SwissFEL} and AWA teams for contributing to the
  formulation of optimization problems. 
  We gratefully acknowledge the computing resources provided on Bebop,
  a high performance computing cluster operated by the 
  Labratory Computing Resource Center at Argonne National Laboratory.
  This work was partly supported by the 
  U.S. Department of Energy, Office of Science, under 
  contract number DE-AC02-06CH11357 and grant number DE-SC0015479. 
\appendix

### 2018-05-20 paper/theor-experiment.tex
small changes
adding DOE in acknowledgments

%Currently we run the simulations on a low resolution space charge grid ($8
  %\times 8 \times 8$) as a proof of concept.
%In a next step a simple Python script would be applied to iteratively rerun
  %the optimizer with increasing space-charge resolution when the limits of the
  %design variables are narrow enough (iteratively diminishing search space). 
%\vspace{-7em}
\subsection{AWA Photoinjector Optimization} \label{awaproblem}
Next we apply the optimization framework to the high charge beam line at the AWA, 
which consists of an rf photocathode gun, two solenoids, six linear accelerating cavities, 
followed by four quadrupoles and a stripline kicker, as shown in Fig. \ref{awa-pic}. 
Portions of this beam line served as the injector for Two Beam 
Acceleration (TBA) experiments in the past [X1], and will be used for 
similar experiments in the future [X2]. Main constraints on prior experimental 
results were the beam size and emittance as the beam passed through small aperture 
wakefield structures located after the quadrupoles. 
In an attempt to maximize charge transmission in upcoming experiments
we optimize several parameters in the gun and quadrupole leading into the 
TBA section of the beam line.
%\vspace{-1em}

@@

\vspace{-1em}
We also chose this model to demonstrate the frameworks ability to tackle larger problems.
The number of design variables is eight, and the number of objectives is seven.
This problem is also less predictable, due to the high dimensionality 
and nonlinear effects such as space charge. There is not a definite idea
before hand where the Pareto front will fall. This work is also meaningful in that
it will guide future operations at the AWA.

@@

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours,
location of Pareto fronts at $s_1$ and $s_2$, and the design variables at the Pareto front. From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated about 30 to 40 fewer generations than the other experiments.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto1}, we can see this is not the case.
The Pareto fronts at $s_1$, are nearly identical. We could expect
they would continue to converge given more time. 
When looking at the Pareto front for $s_2$, we see ex-4 has a slight 
edge over the others.  
\begin{figure}
	\centering
	\begin{subfigure}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto1}
		\caption{Pareto fronts at $s_1$.}
		\label{exparetos1}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
	\centering
	\includegraphics[width=\textwidth]{Report/ex-pareto2}
	\caption{Pareto fronts at $s_2$.}
	\label{exparetos2}
	\end{subfigure}
	\caption{Comparison of Pareto fronts for initial experiments.}
	\label{expareto}
\end{figure}

>>>>

%Currently we run the simulations on a low resolution space charge grid ($8
  %\times 8 \times 8$) as a proof of concept.
%In a next step a simple Python script would be applied to iteratively rerun
  %the optimizer with increasing space-charge resolution when the limits of the
  %design variables are narrow enough (iteratively diminishing search space). 
%\vspace{-7em}
\subsection{AWA Photoinjector Optimization} \label{awaproblem}
Next we apply the optimization framework to the high charge beam line at the AWA. 
As shown in Fig.  \ref{awa-pic}, this beam line consists of an rf photocathode gun, 
two solenoids, six linear accelerating cavities, 
followed by four quadrupoles and a stripline kicker. 
This beam line has served as the injector for Two Beam 
Acceleration (TBA) experiments in the past [X1], and will be used for 
similar experiments in the future [X2]. Main constraints on prior experimental 
results were the beam size and emittance as the beam passed through small aperture 
wakefield structures located after the quadrupoles. 
In an attempt to maximize charge transmission in upcoming experiments
we optimize several parameters in the gun and quadrupole leading into the 
TBA section of the beam line.
%\vspace{-1em}

@@

%\vspace{-1em}
We also chose this model to demonstrate the frameworks ability to tackle larger problems.
The number of design variables is eight, and the number of objectives is seven.
This problem is also less predictable, due to the high dimensionality 
and nonlinear effects such as space charge. There is not a definite idea
before hand where the Pareto front will fall. This work is also meaningful in that
it will guide future operations at the AWA.

@@

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours,
location of Pareto fronts at $s_1$ and $s_2$, and the design variables at the Pareto front. From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated about 30 to 40 fewer generations than the other experiments.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto}, we can see this is not the case.
The Pareto fronts at $s_1$, are nearly identical. We could expect
they would continue to converge given more time. 
When looking at the Pareto front for $s_2$, we see ex-4 has a slight 
edge over the others.  
\begin{figure}
	\centering
	\begin{subfigure}{0.49\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto1}
		\caption{Pareto fronts at $s_1$.}
		\label{expareto1}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
	\centering
	\includegraphics[width=\textwidth]{Report/ex-pareto2}
	\caption{Pareto fronts at $s_2$.}
	\label{expareto2}
	\end{subfigure}
	\caption{Comparison of Pareto fronts for initial optimization experiments.}
	\label{expareto}
\end{figure}

Next compare the design variables at each point on the 
Pareto front.
\begin{figure}
	\centering
	\begin{subfigure}{1\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto1}
		\caption{Pareto fronts at $s_1$.}
		\label{bounds1}
	\end{subfigure}
	\begin{subfigure}{1\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Report/ex-pareto2}
		\caption{Pareto fronts at $s_2$.}
		\label{bounds2}
	\end{subfigure}
	\caption{Comparison of design variables for each Pareto front.}
	\label{bounds}
\end{figure}
